---
title: "My First Blog"
author: "Jiale He"
date: "10/28/2019"
output: html_document
---

In Data Science I, we've learned multiple techniques dealing with data. For instance, ggplot2 and tidyverse are very commonly used packages for data visualization.
In the midterm project, we were requested to finish a data visualization job with the shiny app. The datasets are about the environmental impacts of different cuts of beef. It takes me a very long time to figure out a better way other than bar plot: 2D density plot.

In this midterm project, I used learned techniques and knowledge, but also new stuff from self-study. 
Now let's talk about my masterpiece!

It might first come in mind that how could 2D density plot show the environmental impact? It actually does and from my perspective, pretty nicely done!

First of all, it is to deal with the data.
The dataset is called carcass, and I cleaned the data a bit for easier usage later.
```{r}
carcass<- read.csv('file/carcass_clean.csv')
head(carcass,10)
```

Because I never find a package that has already implemented with an interactive cow map, so some of the steps are done manually. To get a cow map, I have to first put the cow map in the plot. I checked online a lot and found that the 2D density plot can be easily applied on geographic maps with build-in longitude and latitude values.

All you have to do is find the location, give it a value, Woa! you have a nice density plot. However, there's no pre-existed cow map package. 

So to plot a 2D density plot on a cow, I have to find another way round. First, is to have the cow map appear in R. I did not find a way to "plot" the cow, but to either draw the cow or put the cow into the plot as background.

I read in the picture use the png package, and plot with cowplot package. Unfortunately, it is just a picture and I could do nothing with it. As you can see, the background is white just, if you are familiar with ggplot2, the default plot background is always grey.
```{r message=FALSE}
library(png)
library(cowplot)
im <- readPNG("file/cow.png")
beefcut <- ggdraw()+
  draw_image(im)
beefcut
```

I used the other option: I made the cow map a background of my ggplot, using grid package. And it worked! The background is grey and it's in a plot!
```{r message=FALSE}
library(grid)
library(ggplot2)
ggplot()+
  annotation_custom(rasterGrob(im,width = unit(1,"npc"), height = unit(1,"npc")), -Inf, Inf, -Inf, Inf)
```

So I solved the picture issue, what's next? Here comes the manmual part. I set the x and y scale from 0 to 100 for this background, and worked with my teammates with locating all the cuts into file vec.csv.
```{r}
ggplot()+
  annotation_custom(rasterGrob(im,width = unit(1,"npc"), height = unit(1,"npc")), -Inf, Inf, -Inf, Inf)+
  scale_x_continuous(expand=c(0,0), lim=c(0,100)) +
  scale_y_continuous(expand=c(0,0), lim=c(0,100))
vec<-read.csv("file/vec.csv")
head(vec,10)
```

As I mentioned, the cow map is not a geographic map, and there's no cow-related package for me to use. To plot the 2D density plot on this blank plot (the cow map is just a background), I need to create points on the districts that we manually measured. To begin with, the carcass_clean.csv dataset needs further change. The packages we learned such as tidyverse are really useful here. Tidyr and pipe within tidyverse are used below.

```{r message=FALSE}
library(tidyverse)
carcass2<- 
  # alter the original environment impacts to per pound of beef cut and chage units.
  # %>% sign is pipe, means next step, straight forward logic.
carcass %>% 
  # mutate adds new columns with the value you want
  mutate(water.1000gal.lb = water.rate/total_weight*1000,co2e.kg.lb = co2.emision/total_weight*1000*0.454,landuse.bs.lb = (land.use/total_weight*4047)/420) %>% 
  # select displays and gives the colums you want
  select(cut, total_weight, protein.lb, Kcal.lb,water.1000gal.lb,co2e.kg.lb,landuse.bs.lb)
carcass4 <- carcass2 %>% 
  # magnify weight value for better density plot
  mutate(tot_weight_10 = total_weight*10, tot_weight_100 = total_weight*100)
head(carcass4,10)
```

As shown above, water.1000gal.lb means 1000gallons of water usage per pound of the cut of beef, so does the co2 emissions, land usage columns, etc. These values will then become the number of points randomly generated on the blank plot. Geom_density2d can accept these generated points. For example, Round-cut uses 99.29 thousand gallons of water per pound, recall the location values of round-cut, 41.60    62.5    63.0    87.0 will give us a square on the cow map, 99 points will be randomly generated in this square. To generate all these points, I have to write a function within the function. How to write a function and using for loop are important topics we covered in classes.

```{r}
# The first function rwloc is used to calculate weight, it generate random location values(newx,newy) within the square I restricted (x1,x2,y1,y2)
rwloc <- function(cutname,x1,x2,y1,y2){
  # cutname is going to be the cutnames from vec.csv, x1 x2 y1 y2 are the corners of the square we manually measured.
  c4<- carcass4[rep(rownames(carcass4),carcass4$tot_weight_100),]
  c5<- c4 %>% 
    # pipe and tidyr again! filter select coditions you set.
    filter(cut == as.character(cutname))
  c5$newx <- sample(x1:x2,nrow(c5),replace = TRUE)
  c5$newy <- sample(y1:y2,nrow(c5),replace = TRUE)
  return(c5)
}
# The second function utilize the first function to repeat this step for the value of weight
manual <- function(vec){
  # create empty data.frame
  a<- data.frame()
  for (i in 1:nrow(vec)){
    b<- rwloc(vec[i,1],vec[i,2],vec[i,3],vec[i,4],vec[i,5])
    # rbind here adds new rows into the dataframe.
    a<- rbind(a,b)
  }
  return(a)
}
weight <- manual(vec)
weight[1:10,6:11]
nrow(weight)
```
As shown, I generated repeated rows of values but the location value newx and newy are different. The newx and newy will give us randomly located points.

Dataset Weight has 114975 rows, it's all repeated points with differnt location value. So I can plot them using ggplot2 stat_density2d on the cow map.
```{r}
ggplot(data=weight,aes(x = newx, y = newy))+
  annotation_custom(rasterGrob(im,width = unit(1,"npc"), height = unit(1,"npc")), -Inf, Inf, -Inf, Inf)+
  scale_x_continuous(expand=c(0,0), lim=c(0,100)) +
  scale_y_continuous(expand=c(0,0), lim=c(0,100))+
  stat_density2d(alpha = 0.5, geom = "polygon", color = "pink")+
  # the 0 to 100 scale is for us to measure the cuts, as the final product, I don't need them
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  # give my plot a label
  labs(title = "2D Density Plot for Beef Cut Chart")
```

As a result, we could observe that ground beef, blended burger, round part, chuck part and rib part has the highest beef yield. The plot shown contains every cut.

I personally think this project is very meaningful, it drives me to be creative, and this 2d density plot was the only cow plot among all the midterm products. I think it's very unique. It's a little bit sketchy and not good-looking enough, but I utilize the package like tidyr, pipe, ggplot2 and plotly (not included here) learned in class, and techniques such as function, for loops are vital in my project. Apart from the learned knowledge, the project made me learned new packages such as grid, cowplot (actually there are tens more but sadly not useful to be applied in my project). For this project, I am learning and reviewing at the same time, it's fun and challenging but with the product we had, I am more than proud of myself and my teammates Danni and Eddie. The final product can be found [here](https://jialehe3.shinyapps.io/midtermbeef_3/). You can play around with it!




